---
title: "Example plots"
subtitle: Analysis updated on `r Sys.Date()`
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

<br>
<br>
<br>

<!--Initialize-->

``` {r setup, include=FALSE}

### Install necessary packages and get started

library(reshape2)
library(tidyverse)
library(ggplot2)
library(plotly)
library(ggrepel)
library(bindata)
library(odin)
library(fitR)
library(knitr)
library(EasyABC)
library(gridExtra)
library(odin)
library(lubridate)
library(EasyABC)
library(gridExtra)
library(kableExtra)
library(plyr)
library(dplyr)
library(data.table)
library(scales)
library(EasyABC)
library(patchwork)

library(tidyr)
library(readr)
library(purrr)
library(tibble)
library(stringr)
library(forcats)
library(network)
library(tidygraph)
library(ggraph)
library(visNetwork)
library(networkD3)
library(ggmosaic)
library(formattable)
library(DT)
library(reshape)
library(here)
library(fs)
library(hrbrthemes)

library(plotly)

lang_output <- function(x, lang) {
  cat(c(sprintf("```%s", lang), x, "```"), sep = "\n")
}
r_output <- function(x) lang_output(x, "r")

knitr::opts_chunk$set(
  fig.width = 9.5,
  fig.height = 8,
  eval=TRUE,
  echo=FALSE,
  warning=FALSE,
  cache=FALSE,
  message=FALSE,
  include=TRUE
  )
```

```{r}
# navigator.dir = path("/Volumes/GoogleDrive/Shared drives/Disparities Navigator")
# d <- read.csv(here::here(navigator.dir, "LAC_AAIRs_bySiteSex5YearRangeRaceEth_20002019_03FEB22.csv"), na.strings = c("NA", "~"))
# d$csp_race_mod <- recode_factor(d$csp_race_mod, "South Asian (Indian/Pakistani/Sri Lankan/Bangladesh/etc)" = "South Asian")
# d$csp_race_mod <- factor(d$csp_race_mod, levels=c(sort(levels(d$csp_race_mod), decreasing=F)))
# write_rds(d, here::here("data/d.rds"))
d = readRDS(here::here("data/d.rds"))

```


```{r}
# Top N cancer sites
d.top = d %>% filter(SEERCODE != "All Cancer Sites", sex == "Male and Female",csp_race_mod == "All Races/Ethnicities", YEARRANGE == "All (2000-2019)") %>% 
  #group_by("SEERCODE") %>% 
  arrange(desc(aair)) %>% 
  top_n(10, aair) %>% 
  select(SEERCODE) %>% 
  as.matrix() 
```


# Grouped barplot

## Races X years
```{r}

## Grouped barplot groups = races and years + errorbars 
d %>% filter(SEERCODE =="All Cancer Sites", sex == "Male and Female") %>% 
  ggplot(aes(x=reorder(csp_race_mod,aair), y=aair, fill=YEARRANGE)) +
  geom_bar(position=position_dodge(width=.8), stat="identity", width=.7, colour='black') +
  geom_errorbar(aes(ymin=lclair, ymax=uclair), width=.2,position=position_dodge(.8)) +
  coord_flip() + 
  theme_ipsum(axis_text_size = 15, axis_title_size = 15)

```

## Races X years
```{r}

## Grouped barplot groups = races and years + errorbars 
d %>% filter(SEERCODE %in% d.top, sex == "Male and Female", YEARRANGE=="All (2000-2019)") %>% 
  ggplot(aes(x=reorder(csp_race_mod,aair), y=aair, fill=SEERCODE)) +
  geom_bar(position=position_dodge(width=.8), stat="identity", width=.7, colour='black') +
  geom_errorbar(aes(ymin=lclair, ymax=uclair), width=.2,position=position_dodge(.8)) +
  coord_flip() + 
  theme_ipsum(axis_text_size = 15, axis_title_size = 15)



    d.all = d %>% filter(SEERCODE %in% d.top, sex == "Male and Female", YEARRANGE == "All (2000-2019)") %>% 
      filter(csp_race_mod %in% c("All Races/Ethnicities", "Hispanic White", "Non-Hispanic White", "Black"))
    
    ggplot(d.all, aes(x=(csp_race_mod), y=aair, fill=SEERCODE)) +
      geom_bar(position=position_dodge(width=.8), stat="identity", width=.7, colour='black') +
      geom_errorbar(aes(ymin=lclair, ymax=uclair), width=.2,position=position_dodge(.8)) +
      coord_flip() + 
      theme_ipsum(axis_text_size = 15, axis_title_size = 15)

```


# Single barplot

## Top N sites, all races

```{r}
## Top N sites -- all races
d %>% filter(SEERCODE %in% d.top, sex == "Male and Female", csp_race_mod == "All Races/Ethnicities", YEARRANGE == "All (2000-2019)") %>% 
  ggplot(aes(x = reorder(SEERCODE, aair), y=aair, fill=SEERCODE)) +
  geom_bar(position=position_dodge(width=.8), stat="identity", width=.5, colour='black') +
  geom_errorbar(aes(ymin=lclair, ymax=uclair), width=.2,position=position_dodge(.8)) +  
  coord_flip() +
  theme_ipsum(axis_text_size = 15, axis_title_size = 15)
```

## Top N sites, Black ethnicity
```{r}
d %>% filter(SEERCODE %in% d.top, sex == "Male and Female", csp_race_mod == "Black", YEARRANGE == "All (2000-2019)") %>% 
  ggplot(aes(x = reorder(SEERCODE, aair), y=aair, fill=SEERCODE)) +
  geom_bar(position=position_dodge(width=.8), stat="identity", width=.5, colour='black') +
  geom_errorbar(aes(ymin=lclair, ymax=uclair), width=.2,position=position_dodge(.8)) +  
  coord_flip() + 
  #labs(title = "Race/Ethnicity: Black, Top 10 Cancers") + 
  theme_ipsum(axis_text_size = 15, axis_title_size = 15)
```

## Races for a specific cancer site: Breast
```{r}
d %>% filter(SEERCODE == "Breast", sex == "Male and Female", csp_race_mod != "All Races/Ethnicities", YEARRANGE == "All (2000-2019)") %>% 
  ggplot(aes(x = reorder(csp_race_mod, aair), y=aair, fill=csp_race_mod)) +
  geom_bar(position=position_dodge(width=.8), stat="identity", width=.5, colour='black') +
  geom_errorbar(aes(ymin=lclair, ymax=uclair), width=.2,position=position_dodge(.8)) +  
  coord_flip() + 
  theme_ipsum(axis_text_size = 15,
              axis_title_size = 15)
```

## Races for a specific cancer site: Prostate
```{r}
d %>% filter(SEERCODE == "Prostate", sex == "Male and Female", csp_race_mod != "All Races/Ethnicities", YEARRANGE == "All (2000-2019)") %>% 
  ggplot(aes(x = reorder(csp_race_mod, aair), y=aair, fill=csp_race_mod)) +
  geom_bar(position=position_dodge(width=.8), stat="identity", width=.5, colour='black') +
  geom_errorbar(aes(ymin=lclair, ymax=uclair), width=.2,position=position_dodge(.8)) +  
  coord_flip() +
  theme_ipsum(axis_text_size = 15, axis_title_size = 15)
```


# Faceted barplots

## Top N sites, facet races
```{r}
d %>% filter(SEERCODE %in% d.top, sex == "Male and Female", YEARRANGE == "All (2000-2019)") %>% 
  ggplot(aes(x = reorder(SEERCODE, aair), y=aair, fill=SEERCODE)) +
  geom_bar(position=position_dodge(width=.8), stat="identity", width=.7, colour='black') +
  geom_errorbar(aes(ymin=lclair, ymax=uclair), width=.2,position=position_dodge(.8)) +  
  facet_wrap(~csp_race_mod) +
  theme_ipsum(axis_text_size = 12, axis_title_size = 12) +
  coord_flip()

# # Facetwrap (don't like this)
# ggplot(test, aes(fill=csp_race_mod, y=aair, x=csp_race_mod)) + 
#   geom_bar(position="dodge", stat="identity") +
#   scale_fill_viridis(discrete = T, option = "E") +
#   ggtitle("Studying 4 species..") +
#   facet_wrap(~SEERCODE) +
#   theme_ipsum(axis_text_size = 15, axis_title_size = 15) +
#   theme(legend.position="none") +
#   xlab("")
```

## Top N sites, facet by site
```{r, fig.height=10, fig.width=12}
#test = 
  d %>% filter(SEERCODE %in% d.top, sex == "Male and Female", YEARRANGE == "All (2000-2019)") %>% 
  ggplot(aes(x = reorder(csp_race_mod, aair), y=aair, fill=csp_race_mod)) +
  geom_bar(position=position_dodge(width=.8), stat="identity", width=.7, colour='black') +
  geom_errorbar(aes(ymin=lclair, ymax=uclair), width=.2,position=position_dodge(.8)) +  
  facet_wrap(~SEERCODE) +
  coord_flip() +
  theme_ipsum(axis_text_size = 12, axis_title_size = 12)
#ggplotly(test)
```




# Line plots

## Races over time
```{r}
d.all = d %>% filter(SEERCODE == "All Cancer Sites", sex == "Male and Female", YEARRANGE != "All (2000-2019)")
g1 = ggplot(d.all, aes(x = YEARRANGE, y = aair, group=csp_race_mod, color = csp_race_mod,
                           text=paste( 
                             paste("Cancer site: ", SEERCODE),
                             paste("Race/ethnicity: ", csp_race_mod),
                             paste("Cases: ", cases, sep=""), 
                                       paste(" Age-adjusted IR (95% CI): ", aair, " (",lclair, ", ",uclair,")", sep=""),sep = "<br>"))) + 
  geom_line(size = .5) + geom_point(size = 1.5, alpha = 0.5) + 
  labs(color = "User-Selected Races", title = paste0("All Cancer Sites", " Over Time, User Selected Races"), x = "Year Range", y = "Age-Adjusted Incidence Rate") +
  #geom_errorbar(aes(ymin=lclair, ymax=uclair), width=2,position=position_dodge(0.1)) +
  theme_ipsum(axis_text_size = 15, axis_title_size = 15)
g2<-ggplotly(g1, tooltip = "text")
g2
```



## With trendline

```{r}

d.all = d %>% filter(SEERCODE == "All Cancer Sites", sex == "Male and Female", YEARRANGE != "All (2000-2019)")
d.all$yearNum = d.all$YEARRANGE
d.all$yearNum = unclass(d.all$yearNum)
# d.all %>% 
#   plot_ly(x=~yearNum, y=~aair, type = 'scatter', mode='lines', color=~csp_race_mod)
# 
# fit <- lm(aair ~ yearNum, data = d.all)

d2 = data.table(d.all)
d3 = d2[,as.list(coef(lm(aair~yearNum))),by=csp_race_mod]
d4 = merge(d2, d3, by="csp_race_mod")
d4 = data.frame(d4)

g1 = ggplot(d4, aes(x = YEARRANGE, y = aair, group=csp_race_mod, color = csp_race_mod,
                           text=paste( 
                             paste("Cancer site: ", SEERCODE),
                             paste("Race/ethnicity: ", csp_race_mod),
                             paste("Cases: ", cases, sep=""), 
                             paste(" Age-adjusted IR (95% CI): ", aair, " (",lclair, ", ",uclair,")", sep=""),
                             paste("Trend: ", round(yearNum.y/5,2), "units change in rate per year" ) ,
                             sep = "<br>"))) + 
  geom_smooth(aes(YEARRANGE, aair), method = "lm", se = F, colour =  "black", size = .2) +
  geom_point(size = .5) + geom_point(size = 1.5, alpha = 0.5) + 
  labs(color = "User-Selected Races", title = paste0("All Cancer Sites", " Over Time, User Selected Races"), x = "Year Range", y = "Age-Adjusted Incidence Rate") +
  #geom_errorbar(aes(ymin=lclair, ymax=uclair), width=2,position=position_dodge(0.1)) +
  theme_ipsum(axis_text_size = 15, axis_title_size = 15)
g1

g2<-ggplotly(g1, tooltip = c("text")) 
g2


```












## Races over time, specific cancer site: Breast
```{r}
d.all = d %>% filter(SEERCODE == "Breast", sex == "Female", YEARRANGE != "All (2000-2019)")
g1 = ggplot(d.all, aes(x = YEARRANGE, y = aair, group=csp_race_mod, color = csp_race_mod,
                       text=paste( 
                         paste("Cancer site: ", SEERCODE),
                         paste("Race/ethnicity: ", csp_race_mod),
                         paste("Cases: ", cases, sep=""), 
                         paste(" Age-adjusted IR (95% CI): ", aair, " (",lclair, ", ",uclair,")", sep=""),sep = "<br>"))) + 
  geom_line(size = .5) + geom_point(size = 1.5, alpha = 0.5) + 
  #geom_errorbar(aes(ymin=lclair, ymax=uclair), width=2,position=position_dodge(0.1)) +
  #scale_color_viridis(discrete = TRUE, option="turbo") +
  theme_ipsum(axis_text_size = 15, axis_title_size = 15)
g2<-ggplotly(g1, tooltip = "text")
g2
```


## Top 10 sites over time
```{r}
g1 = d %>% filter(sex == "Male and Female",csp_race_mod == "All Races/Ethnicities", SEERCODE %in% as.matrix(d.top), YEARRANGE != "All (2000-2019)") %>% 
  ggplot(aes(x = YEARRANGE, y = aair, group=SEERCODE, color = SEERCODE,
                       text=paste( 
                         paste("Cancer site: ", SEERCODE),
                         paste("Race/ethnicity: ", csp_race_mod),
                         paste("Cases: ", cases, sep=""), 
                         paste(" Age-adjusted IR (95% CI): ", aair, " (",lclair, ", ",uclair,")", sep=""),sep = "<br>"))) + 
  geom_line(size = .5) + geom_point(size = 1.5, alpha = 0.5) + 
  geom_errorbar(aes(ymin=lclair, ymax=uclair), width=2,position=position_dodge(0.1)) +
  #scale_color_viridis(discrete = TRUE, option="turbo") +
  theme_ipsum(axis_text_size = 15, axis_title_size = 15)
ggplotly(g1, tooltip = "text")

```





# Heatmaps

## Races X years 
```{r}
dmat = d %>% filter(SEERCODE == "All Cancer Sites", sex == "Male and Female") %>% select(c(YEARRANGE, csp_race_mod, aair)) %>% pivot_wider(names_from = YEARRANGE, values_from = aair) %>% as.data.frame()
rownames(dmat) = dmat$csp_race_mod
dmat$csp_race_mod = NULL
dmat = as.matrix(dmat)

plot_ly(x=colnames(dmat), y=rownames(dmat),
        z=~dmat,
        type="heatmap",
        showscale=T)
```


## Top N sites x races
```{r}
dmat = d %>% filter(YEARRANGE == "All (2000-2019)", sex == "Male and Female", SEERCODE %in% d.top) %>% 
  select(c(SEERCODE, csp_race_mod, aair)) %>% 
  filter(SEERCODE!="All Cancer Sites") %>% 
  pivot_wider(names_from = csp_race_mod, values_from = aair) %>% 
  as.data.frame()
rownames(dmat) = dmat$SEERCODE
dmat$SEERCODE = NULL
dmat = as.matrix(dmat)

plot_ly(x=colnames(dmat), y=rownames(dmat),
        z=~dmat,
        type="heatmap",
        showscale=T)
```








